---
title: "Discussion 1"
author: "Ismam Ferdous"
date: "`r Sys.Date()`"
output: html_document
---

<style>h1 b{font-weight:normal;font-size:medium;font-style:italic}</style>

```{r setup, include=FALSE}
# this is called a "setup" chunk and is often found at the top of Rmd files
# setting "include=FALSE" let's it run but completely hides all code + output

# this line below sets default chunk options for the entire document
knitr::opts_chunk$set(echo=TRUE,fig.width=5,fig.height=4,fig.align="center")
```


# Question 1  <b>(R/Rstudio check)</b>

Have you followed instructions from lecture and installed the right R version, installed `tidyverse`, as well as setup and organized your files?

```{r}
# no written response needed, simply knit the document and
# check the output, this chunk should print:
#  1. the R version currently installed,
#  2. the tidyverse import messages, and
#  3. the file folder structure of your STAT240 class folders
# (remember to set your working directory if running this in console!)

sessionInfo()
library(tidyverse)
list.dirs("../../",recursive = TRUE)
```


# Question 2 <b>(basic R practice)</b>

The following dataset comes from openintro.org. 3 different bodies of water in Bimini Lagoon, Bahamas were tested and their salinity in parts per thousand measured repeatedly over a period of time. The data are shown below:

    Site 1: 37.54,37.01,36.71,37.03,37.32,37.01,37.03,37.7,37.36,36.75,37.45,38.85 
    Site 2: 40.17,40.8,39.76,39.7,40.79,40.44,39.79,39.38     
    Site 3: 39.04,39.21,39.05,38.24,38.53,38.71,38.89,38.66,38.51,40.58    

For each question part below, insert a code chunk, do the necessary R work, then write your final response on a new line starting with "> RESPONSE: " (similar to question 1 above).


## Part A

For each site, create a data vector storing the salinity measurements for that site. Print each vector out to show they have been created correctly.


```{r}
site_1 <- c(37.54,37.01,36.71,37.03,37.32,37.01,37.03,37.7,37.36,36.75,37.45,38.85)
site_2 <- c(40.17,40.8,39.76,39.7,40.79,40.44,39.79,39.38)
site_3 <- c(39.04,39.21,39.05,38.24,38.53,38.71,38.89,38.66,38.51,40.58)

site_1
print("> RESPONSE: The mean salinity for Site 1 is approximately 37.31.")

site_2
print("> RESPONSE: The mean salinity for Site 2 is approximately 40.10.")

site_3
print("> RESPONSE: The mean salinity for Site 3 is approximately 38.92.")



```


## Part B

For each site, compute and print the following values: sample size (i.e. number of observations), mean salinity, median salinity, min and max values, the 90-th percentile, and the standard deviation.

Also answer: Which site had the highest mean? Which site had the greatest range (i.e. max vs min) of values?


```{r}
    length(site_1)
    print("> RESPONSE: The sample size of site 1 is 12.")

    length(site_2)
         print("> RESPONSE: The sample size of site 2 is 8.")

    length(site_3)
          print("> RESPONSE: The sample size of site 3 is 10.")

    
    #mean salinity for each site : 
    
    mean(site_1)
     print("> RESPONSE: The mean for site 1 is 37.31.")
    mean(site_2)
    print("> RESPONSE: The mean for site 2 is 40.10.")
    mean(site_3)
    print("> RESPONSE: The mean for site 3 is 38.94.")

    #finding the median for each site : 
    
    median(site_1)
         print("> RESPONSE: The median for site 1 is 37.17.")

    median(site_2)
             print("> RESPONSE: The median for site 2 is 39.98.")

    median(site_3)
             print("> RESPONSE: The median for site 3 is 38.8.")

    
    # finding the minimum and maximum for each site : 
    
    min(site_1)
    max(site_1)
    
                 print("> RESPONSE: The min is 36.71 and the max is 38.85 for site 1")


    min(site_2)
    max(site_2)
    
     print("> RESPONSE: The min is 39.38 and the max is 40.80 for site 2")

    min(site_3)
    max(site_3)
    
     print("> RESPONSE: The min is 38.24 and the max is 40.58 for site 3")
    
    #finding the 90th percentile :
    #we use the quantile function, and then we specify with a decimal value
    quantile(site_1, probs = 0.9)
    quantile(site_2, probs = 0.9)
    quantile(site_3, probs = 0.9) 
    
     print("> RESPONSE: 90th PCT for site 1: 37.684. Site 2: 40.79. Site 3: 39.34.")
    
    # finding the standard deviation : 
    
    sd(site_1)
    sd(site_2)
    sd(site_3)
    
    
    
    print("> RESPONSE: Standard devition for site 1 : 0.57. Site 2: 0.53. Site 3: 0.64")
    
   
    
    
    print("Site 2 had the highest mean, site 2 also had the greatest range")
    


```




## Part C

A [boxplot](https://en.wikipedia.org/wiki/Box_plot) is a simple plot for visualizing one set of values or comparing several different sets. It can be easily made with the `boxplot()` function by giving it each data vector as inputs, e.g. `boxplot(x,y,z,...)` where x,y,z,... are data vectors.

Make a simple boxplot showing the salinity measurements from each site. For now, don't worry about titles or axes labels.

```{r}

boxplot(site_1, site_2, site_3)


```





