---
title: "Discussion 7"
author: "Ismam Ferdous"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,fig.width=5,fig.height=4,fig.align="center")

###
### REMEMBER TO:  1. set your working directory
###               2. import necessary packages (run line below)

library(tidyverse)
```


## Review notes

As usual please review the intro to probability sections we covered this past week. Make sure you understand each of the following key concepts:

 - Random variables
 - Outcomes, events, and sample space
 - Axioms of probability
 - Intersections and unions of events
 - Mutual exclusivity
 - Discrete and continuous distributions
 - PMFs and PDFs, and how they differ

## Probability practice

1. Give an example each of a discrete and continuous random variable For each of your examples of a discrete and continuous RV, can you give an example PMF and PDF that could plausibly be its distribution function?

```{r}

# a discrete random heads is one that takes on countable values. An example of this could be the the # of heads when flipping 3 coins. 

# Random variables : X = number of heads
# Possible values : 
# PMF : probability mass function, gives the probability of each value. 

x <- 0:3  # possible outcomes
pmf <- c(1/8, 3/8, 3/8, 1/8) # there are 8 total outcomes. 

discrete_rv <- data.frame(X = x, PMF = pmf)
print(discrete_rv)


```


2. For each of the following statements, think about them carefully and respond briefly to each.

   a. Why do we need the axioms of probability?
   b. What are "outcomes" vs "events"?
   c. Can you ever observe something outside the sample space?
   d. What does mutually exclusive mean?
   e. Why do we have to distinguish discrete vs continuous variables?
   f. What ways can you specify the distribution of an RV? Is there a best way?
   g. What does the PMF tell you for each outcome k?
   h. What does the PDF tell you for each outcome k an how does this differ from PMFs?
   i. What is the empirical rule and where does it apply?
   j. What is an expected value?


```{r}

# a. they give us the basic rules to make probability consistent and logical.

# b. outcomes are a single result (flipping H) and event is a set out outcomes, (getting atleast one head)

# c. no, everything must be part of the sample space by definition

#d. it means 2 events cannot happen at the same time. 

#e. because they behave differently and need different probability tools 

#f. we can do this by doing PMF/PDF, tables, formulas, or graphs, there is no best way, just in the context.

#g. it tells us the probability of that random variable k. 

#h. PDF gives us the density, not exact probability. probability comes from the area under the curve.
#PMFs give exact probabilities for discrete outcomes.

#i. In a normal distribution: ~68% within 1 SD, 95% within 2 SD, 99.7% within 3 SD.

#j. The weighted average or long term mean of a random variable.





```

