---
title: "Discussion 8"
author: "Ismam Ferdous"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,fig.width=5,fig.height=4,fig.align="center")

###
### REMEMBER TO:  1. set your working directory
###               2. import necessary packages (run line below)

library(tidyverse)
```


## Review notes

As usual please review the intro to probability sections we covered this past week. Make sure you understand each of the following key concepts:

 - Expected value
 - Variance & standard deviation
 - Binomial distribution:
   - assumptions
   - parameters
   - PMF function (including factorial & choose)
   - E, Var, SD for binomial
   - *binom functions

## Probability practice

1. Give 2 examples of possible applications of the binomial variable.

```{r}

# 1.  A Drug trial: Counting how many out of 20 patients experience a side effect (yes/no) when taking a new medication

# 2 :  Counting the number of defective items in a batch of electronics where each has a fixed chance of being defective such as electronics. 


```


2. List some ways the binomial assumptions can fail to hold as a model.


```{r}


# 2. Ways Binomial Assumptions Can Fail

# The probability of “success” changes from trial to trial, it is not always constant. 
#Outcomes aren’t truly independent, it effects the others
# There are more than two possible outcomes per trial.


```


3. For each of the following statements, think about them carefully and respond briefly to each.

   a. What is the meaning of an expected value of a variable? How about the EV of a function of a variable?
   b. What is the meaning of the variance of a variable? How about the SD?
   c. Why is the SD often more commonly used instead of the variance to describe spread?
   d. Which assumptions of a binomial are often satisfied more "approximately" than "exactly", and why?
   e. What is the support (i.e. set of possible outcomes with nonzero probability) of a binomial RV?
   f. What does each term of the binomial PMF mean?
   g. Under what conditions will a binomial RV be symmetric about the average? (hint: think about the terms in the PMF).
   
   
```{r}
#3a.

# The expected value of a variable is its average value if you repeat the experiment many times. 
# - The EV of a function of the variable is the average value you get after applying that function to the outcomes.


```
   
   
