---
title: "Discussion 9"
author: "Ismam Ferdous"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,fig.width=5,fig.height=4,fig.align="center")

###
### REMEMBER TO:  1. set your working directory
###               2. import necessary packages (run line below)

library(tidyverse)
```


## Review notes

As usual please review the probability & inference sections we covered this past week. Make sure you understand each of the following key concepts:

 - CDFs and inverse CDFs
 - Normal functions:
   - Parameters
   - PDF function form
   - E, Var, SD
   - *norm functions
   - QQ plots
   - Z-scores
 - Normal approximation of binomial:
   - Condition for usage
   - How to apply
   - Continuity correction
 - Roughly what the LLN means
 - Formula and interpretation of confidence intervals

## Probability practice

1. Give 2 examples of possible normal distributions
```{r}

# Heights of adult men
# Birth weights of babies 

```


2. Practice plotting the normal PDF, CDF, and inverse CDF functions (you can borrow code from lecture)

```{r}

x <- seq(-4, 4, length=100)

# PDF plot
plot(x, dnorm(x), type='l', main='Normal PDF', ylab='Density', xlab='x')

# CDF plot
plot(x, pnorm(x), type='l', main='Normal CDF', ylab='Probability', xlab='x')

# Inverse CDF plot
p <- seq(0, 1, length=100)
plot(p, qnorm(p), type='l', main='Normal Inverse CDF', ylab='Quantile', xlab='Probability')



```


3. For each of the following statements, think about them carefully and respond briefly to each.

   a. What is the support (i.e. set of possible outcomes with nonzero probability) of a normal RV?
   b. What does the empirical rule say?
   c. What are the EV and SD of a normal?
   d. How do you read a QQ plot, what do you look for?
   e. What does standard normal mean?
   f. What's the formula for computing a Z-score of a point?
   g. When and how can you approximate a binomial with a normal?
   h. How do you tell which direction to ±0.5 when applying continuity correction?
   i. What's the general form of the formula for a confidence intervaL?
   j. What exactly does "95% confidence" mean?
