---
title: "Homework 1, Stat 240"
author: "Ismam Ferdous"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```
# Assignment Policy
# - Assignments will be posted regularly (specify day if given, e.g. weekly)
- Assignments are typically posted on Fridays and due weekly. 
# - Due dates: (Due every Friday.)
# - Late policy: (No late submissions will be accepted.)
# - Number of drops allowed: (Lowest 2 discussion worksheets and lowest 2 homework scores will be dropped.)

# Exam Scheduling
# - Number of exams: (2 exams total)
# - Exam dates, times, and locations: (Midterm: Friday, Oct 17 (in-class, regular lecture time/place. 
  -  Final: Tuesday, Dec 16, 7:45–9:45 AM. )
# - Conflicts: Confirm you have checked for conflicts. 
#   If yes, did you contact the instructor? (Yes/No)
# - Emergencies/conflicts: Agree to contact the instructor ASAP with documentation to reschedule.
#   (Failure to do so may result in a 0)

# Final Grades
# - Final grade computation: 

- Components:  
  - Discussion worksheets: 5% (lowest 2 dropped)  
  - Homeworks: 20% (lowest 2 dropped)  
  - Project: 20%  
  - Midterm: 25%  
  - Final: 30%  

# - Agreement: You agree to be evaluated according to the provided breakdown/scale.
Agree.

```
```

# PART B

```{r}

#a.  What is the square root of the base 10 log of 2025?
sqrt(log10(2025))

# b. What are the remainders of 2025 when divided by each of the numbers 2 to 9? 
divisors <- 2:9
remainders <- 2025 %% divisors
data.frame(Divisor = divisors, Remainder = remainders)

# c. What's the correlation coefficient between these points: (3.63,53.1), (3.02,49.7), (3.82,48.4), (3.42,54.2), (3.59,54.9)?
x <- c(3.63, 3.02, 3.82, 3.42, 3.59)
y <- c(53.1, 49.7, 48.4, 54.2, 54.9)
correlation <- cor(x, y)
correlation

# d. How does R handle 0 to the power of 0, and how does this compare with other programs?

0^0  

# In other programs it is also similar, as in Python, Matlb and Java it also returns 1 as well. 



#e.Other than the examples in section 2.3.4, can you come up with another way of producing NaN or ±Inf as an output (without using them in the input)?
# Yes there are other ways of prodocuing Nan or Inf such as : 
0/0
Inf - Inf
log(-1)
log10(-1)


# f. Write two mathematical inequalities using both & and | as well as at least a few mathematical operations, one of which evaluates to TRUE and the other to FALSE.

(3 + 5 > 6) & (10 / 2 == 5) | (4^2 < 10)


(2 * 3 > 10) & (7 - 4 == 2) | (5 + 1 < 5)



#2 

# a. Give 3 examples of invalid variable names (you can just write these out instead of putting them in a chunk).


```

#2variable
#name-of-variable
#my variable

```{r}



# b. Find some R package you like and install it 

install.packages("ggplot2", repos = "https://cran.r-project.org")
library(ggplot2)

ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point()
  
#ggplot package designed for creating a wide variety of data visualizations and plots.



# c. Using the built-in help page, look up prop.test( ) and briefly describe what it does, important arguments, and give a simple example of how to use it.

prop.test(x = 55, n = 100, p = 0.5)

# In summary, prop test basically used to test hypotheses about one or more proportions in groups or samples.
#For prop.test don't forget : 

# x: Number of successes (can be a vector/ matrix for multiple groups)

# n: Number of trials (same length as x)

# p: Hypothesized probability of success (default is 0.5)



```

```{r}
#3 

temps <- c(22.4, 24.1, 23.5, 22.9, 23.4, 24.9, 25.3, 23.9, 22.8, 24.2, 25.0, 24.5, 23.7, 23.1)

  mean(temps)

  median(temps)

  sd(temps)

 min(temps)
 
 max(temps)
 
 
 boxplot(temps)

#The boxplot shows the distribution of daily temperatures over the 2-week period. The median temp would be 23.9 C. One observation I made is that there are no outliers, all of the temperatures are within a range which shows us consistent weather conditions. 

#Website to where I found the DataSet : https://www.ncdc.noaa.gov/cdo-web/


```


